---
title: "suicide_rate"
author: "Paul Eudeline"
date: "21/05/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r import, echo=FALSE}
# general data manipulation
library(tidyverse)

# clean visualisation
library(kableExtra)
library(plotly)
```

## Load the dataset

Load the dataset with the `read_csv` function.

```{r}
suicide_data <- read_csv("master.csv")
```

## Quick look at the file strucuture and content

As a first step let’s have a quick look of the dataset with the `summary` function and then an example of 10 random rows.

```{r summary}
suicide_data%>%
  summary()
```


```{r}
suicide_data%>%
  sample_n(10)%>%
  kable()%>% 
  kable_styling()
```

Look if there is any missing values.

```{r}
suicide_data%>%
  is.na()%>%
  colSums()

```

## Visual overview

We will start our exploration by looking at the worldwide suicide rate (maybe use highcharter for nicer interactive plot).

First lets take a look at the worldwide suicide rates.

```{r}
custom_theme <- theme_light()
  
```


```{r}
# Get global suicides
ww_srate <- suicide_data%>%
  select(year, suicides_no, population)%>%
  group_by(year)%>%
  summarise(suicide_capita = round(sum(suicides_no) / sum(population) * 100000, 2))

# Plot
ww_line_plot <- ww_srate%>%
  ggplot(aes(year, suicide_capita)) +
  geom_line() +
  geom_hline(yintercept = mean(ww_srate$suicide_capita), size=0.5, linetype="dotted") +
  geom_point(aes(color=suicide_capita)) +
  labs(title="Worldwide suicide by year", x="Year", y="Suicides per 100k people", subtitle = "1985 - 2016") +
  scale_color_gradient(low="#192a56", high="#fbc531") +
  custom_theme +
  theme(legend.position = "none")

text_info <- paste("Suicides:", ww_srate$suicide_capita, "\nYear:", ww_srate$year)

ggplotly(ww_line_plot)%>%
  style(text=text_info)
```

Now worldwide suicide rate by gender.

```{r}
ww_srate_gender <- suicide_data%>%
  select(year, sex, suicides_no, population)%>%
  group_by(year, sex)%>%
  summarise(suicide_capita = round(sum(suicides_no) / sum(population) * 100000, 2))

# Plot
ww_line_plot <- ww_srate_gender%>%
  ggplot(aes(year, suicide_capita, color=sex, shape=sex)) +
  geom_line() +
  geom_point() +
  labs(title="Worldwide suicide by gender", x="Year", y="Suicides per 100k people", subtitle = "1985 - 2016") +
  custom_theme +
  theme(legend.position = "top")

ggplotly(ww_line_plot)
```

